### 2014-05-17 journal/article.tex
SUmmary table: Agile is hard, coordinating between teams

% ---------------------------------------------------------------------
% -------------- PREAMBLE ---------------------------------------------
% ---------------------------------------------------------------------

%\documentclass[12pt,a4paper,english,oneside]{article}
\documentclass[preprint,authoryear,12pt]{elsarticle}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\usepackage{calc}
\usepackage{natbib}
\usepackage{url}
%\usepackage{listings}
\usepackage{hyphenat}

\usepackage{supertabular,array}
\usepackage{lipsum}
\usepackage{booktabs}
\usepackage{graphicx}

% Strike through
\usepackage[normalem]{ulem}

% Punctuation for references
\bibpunct{[}{]}{;}{n}{,}{,}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

% Multiple bibliographies
%\usepackage{multibib}
%\newcites{gen}{References}
%\usepackage{bibentry}

% ---------------------------------------------------------------------
% -------------- DOCUMENT ---------------------------------------------
% ---------------------------------------------------------------------

\journal{Information and Software Technology}

\begin{document}

% Extra definitions?
\selectlanguage{english}

\begin{frontmatter}

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\title{Adopting agile in the large: motivations, success factors and challenges}

\author{Kim Dikert}
\ead{kim-karol.dikert@aalto.fi}

\author{Maria Paasivaara}
\ead{maria.paasivaara@aalto.fi}

\author{Casper Lassenius}
\ead{casper.lassenius@aalto.fi}

\address{Aalto University, School of Science, Department of Computer Science and Engineering}

% ---------------------------------------------------------------------
% ------------------------- ABSTRACT ----------------------------------
% ---------------------------------------------------------------------

\begin{abstract}

Agile methods have become an appealing alternative for companies striving to
improve their performance, but the methods are originally designed for small and
individual teams. This poses challenges when introducing agile to large
organizations where development teams can not be autonomous. We reviewed 30
experience reports and case studies describing the process of taking agile
methods into use in large organizations. The primary motive for agile adoption
is the need to improve performance. For a successful transformation the entire
organization must be aligned and all organizational units must participate.
Creating an agile community in the organization will help the change to stick.
Typical problems are misunderstanding of the agile model, and difficulties
coordinating multiple agile teams.

\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

agile \sep transformation \sep large scale \sep adopting agile

\end{keyword}

\end{frontmatter}

% ---------------------------------------------------------------------
% --------------------------- TEXT ------------------------------------
% ---------------------------------------------------------------------

\section{Introduction}

--> Introductory paragraph

--> Motivation for research
-- There are no systematic reviews for this
-- A collection of experience reports can be used for creating a theoretical
   base.
--> There is interest in large scale [ref XP13 workshop]

--> Freudenberg, S. and Sharp, H., "The Top 10 Burning Research Questions from
    Practitioners," IEEE Software, pp. 8-9, 2010.

--> Why is large scale software development different from smaller scale?

--> This literature review is intended for researchers as a theroretical
   and proven foundation for the real problems that organizations face in
   transformations.
   This is also intended to be directly usable by practicioners, to serve as a
   base of general advice with strong motivation (although every organization
   has unique needs).

% ---------------------------------------------------------------------
\section{Background}
\label{sec:background}

Lindvall, M., et al., "Agile Software Development in Large Organizations,"
IEEE Computer, vol. 37, pp. 26-34, 2004.

Boehm, B. and Turner, R., Balancing Agility and Discipline: A Guide for the
Perplexed: Addison-Wesley, 2003.
Boehm, B., "Get ready for agile methods, with care," IEEE Computer, vol. 35,
pp. 64 - 69, 2002.
--> Any hints on size?? Maybe size leaning towards discipline?

\subsection{Things that may need to be explained}

* Scrum

* XP

* Definition of sprint

* Self organization and independence of teams -- This is a question relating how
  to organize management.

* Scrum master must not command and control -- Scrum master as manager??

* People centricity -- agile is not for squeezing efficiency -- agile is not for
  fewer career opportunities in a flatter organization.

* Agile is a mindset -- it can not be studied from a book -- need of coaching

% ---------------------------------------------------------------------
\section{Research method}
\label{sec:method}

The goal of the research was to aggregate evidence on how large organizations
transform into an agile way of working. The research was conducted as a
systematic litterature review based on empirical studies.

\subsection{Research questions}

The following research questions were set to guide the research process.

\begin{itemize}

\item
RQ1: Why do large organizations initiate an agile transformation?

\item
RQ2: How do large scale agile transformations usually proceed; do
     transformations exhibit patterns?

\item
RQ3: What are typical successes and challenges in a large scale agile
     transformation process?

\end{itemize}

\subsection{Definition of large scale}
\label{sec:largescale}

A key question in the study was the definition of \emph{large scale}. Large
scale requires additional coordination and makes communication difficult
compared to software development in small teams.

--> Needs to have complexity in software design?? --> What is complexity in software?
    Is ``complexity'' an issue in agile?
    What will bring complexity into software? Not costs or the length of the
    schedule. --> Having many people will bring the complexity in terms of the
    SDM to use.
--> Some people see cost as a definition for large scale, some see code base
    size [Petersen], some see project duration [Bjarnson?]. These are all
    related, but do not relate to any specific problem. You can spend X amount
    of money, but it does not in itself make any difference if the money is
    spent in an agile way.
--> Formal definitions (EU, etc.) look at project costs or company turnover.
    These definitions do not take a standpoint on complexity of the product, so
    they do not matter in the context of SDMs.

--> Ref: the XP conference
--> [Cockburn 2003] Agile is best suitable for teams around 50 people or fewer.
--> Boehm, B. and Turner, R., Balancing Agility and Discipline: A Guide for the
    Perplexed: Addison-Wesley, 2003. --> There might be a 50 person boundary?
    --> See Boehm also for list of

--> The company must already be with N teams or N people, not growing up to the
    required numbers. Small companies in growth are therefore excluded.

\subsection{Research process}

The research was conducted as an application of Kitchenham's
\cite{Kitchenham2007} guidelines for systematic literature review. The selection
of primary studies was done as a keyword based database search and manual
filtering of included studies. The filtering process was executed independently
by two researchers (Dikert, Paasivaara). The selected primary studies were coded
for data extraction and analysis. The entire process was audited and mentored by
a third researcher (Lassenius).

We deviated from Kitchenham's guideline on the part of data extraction. Instead
of using data extraction forms we analyzed the primary studies by coding. This
deviation was made in order to minimize how possible preconceptions would affect
the data extraction step, and to use a specialized software tool that allowed us
to record evidence in as much detail as possible.
% --> Possibility to find many different viewpoinits for the research questions
% --> Lots of items, which would have lead to much diversity in data extraction form

\subsubsection{Inclusion criteria}
\label{sec:inclusioncriteria}

Based on the research questions and intended focus of the research, we defined
three facets to guide inclusion/exclusion decisions: agile software engineering,
focus on transformation, and large scale. The first facet covers primary studies
focusing on software engineering organizations applying or striving to apply
agile methods. The second facet states that included studies must provide
insights relevant to organizational transformation, and specifically to the
research questions. The third facet underlines the particular contribution we
aim to provide with this research, namely the exclusive focus on large
organizations. These three facets were used for inclusion and exclusion in each
study selection step, including search string design, filtering by abstracts,
and full text filtering.

Examples on topics excluded by the first facet are agile manufacturing, and
applying Scrum practices in management boards, as these do not relate to
software engineering. In addition, we required that the organizational
transformation was aimed to introduce agile methods, which excluded other
development methods than agile \cite{Sagesser2013}, and use of agile
methods in other contexts than software development \cite{Hodgkins2007}.

The facet of \emph{organizational transformation} was interpreted so that the
primary study must present insights on the process of transformation. This
was the most important facet, as the pivot of this research is to examine
how the transformations happen. Examples on excluded topics are comparing the
original and agile development methods \cite{Petersen2010}, discussing use of
agile in a large organization but not describing how the new methods were
introduced \cite{Mishra2011}, and merely presenting agile tools in large scale
use \cite{Kim2012}. These sources were excluded as they did not privide insights
on organizational transformation.
% Also [Fitzgerald 2005] on just presenting use of agile methods, no introducing

Transformation and ``scaling up'' are very closely related and in some cases
they are one and the same. For instance, if transformation begins with a pilot
and spreads gradually through the organization, the process can very well be
seen as a ``scaling up'' journey. However, we excluded cases where a initially
small organization was scaled up \cite{Maranzato2012}, and discussions with a
process centered viewpoint without describing organizational change
\cite{Lyon2008}.
% Also [Greening 2010] on process centered viewpoint without transformation

The facet of \emph{large scale} was interpreted as presented in section
\ref{sec:largescale}, having 50 development personel or 6 teams as as the limit
for large scale. In some cases the source presented very vague indicators on
size. For instance, we deemed the case presented by Cloke \cite{Cloke2007} to be
included as there was indications of large scale considerations, although the
size remained unclear. The case presented by Miller and Haddad \cite{Miller2012}
was deemed to be excluded as there was no indications whatsoever on size. A
borderline exclude was the case presented by Tudor \cite{Tudor2006} as there was
no indication of issues indicating large size, althoug there were several
development teams.
To complicate matters, some papers talked about ``the team'' in singular when
referring to the organization \cite{Hodgkins2007}, making it clearly nontrivial
to judge whether an organization was large based on the choice of words of the
author.

A further example on exclusion by the \emph{large scale} facet were cases with
large organizations but only a single team adopting agile \cite{Fulgham2011}.
We considered cases of single teams (although in large organizations) unrelated
to this research as our focus is on transformation of the entire organization.
Also piloting cases that resulted only in single teams using agile were excluded
\cite{Scott2008}.

% Also [Jensen 2003] with only a single team
% Possibly [Mc Cormic 2005] on single team piloting

In addition to the criteria listed above we excluded studies that were not
empirical. We excluded textbooks, studies merely presenting theories, and other
studies that did not include any case organization. We also excluded studies of
benefits or limitations of agile in general. Lastly, we excluded student
experiments as it is implausible to simulate the dynamics of a large
organization in an experiment.

\subsubsection{Preliminary searches}

% --> Why are the preliminary searches done?

Before proceeding with the database search a number of preliminary searches were
performed. The purpose of the preliminary searches was to create a benchmark of
potentially relevant papers that should be matched by the actual database
search. We also wanted to list top ranked hits by trivial keywords that the
more complex final search string might miss. Initial searches were made using
keywords that were as general as possible, including ``agile transformation''
and ``large scale agile''. Additionally, a trial keyword search was done in the
selected databases (listed in table \ref{table:databases}) using the following
search string:

\begin{verbatim}
((agil* OR scrum OR xp OR lean) AND
(transform* OR transit* OR change OR migrat*) AND
((large AND (scale OR organization))) OR enterprise)
\end{verbatim}

From these preliminary searches we compiled a list of 117 possibly relevant
papers that should be matched by the actual database search.

\subsubsection{Study selection}

The study selection process inlcuded three steps, each step refining the set of
selected papers. First a database search was used to gather an initial set of
papers. This set was then refined by filtering out abstracts and finally full
text filtering. The study selection process is outlined in Figure
\ref{fig:selectionprocess}.

\begin{figure}[b]
  \begin{center}
    \includegraphics[width=1\textwidth]{graphics/research_process.pdf}
    \caption{Study selection process}
    \label{fig:selectionprocess}
  \end{center}
\end{figure}

A keyword search in online databases was chosen as the means to find potential
primary studies. The databases included in the search are listed in Table
\ref{table:databases}. All databases supported use of complex boolean logic in
searches which allowed us flexibility when constructing the search strings.

\begin{table}
    \begin{tabular}{ p{0.28\textwidth} l l }
        \toprule
        Database         & URL                         & N of matches   \\
        \midrule
        IEEExplore       & http://ieeexplore.ieee.org      & 745 \\ 
        ACM              & http://dl.acm.org               & 168 \\
        Scopus           & http://www.scopus.com/home.url  & 1596 \\
        Web of Knowledge & http://apps.webofknowledge.com  & 786 \\
        \bottomrule
    \end{tabular}
    \caption{Databases included in search, and number of matched articles}
    \label{table:databases}
\end{table}

We constructed a search string based on the facets \emph{agile software
deveolpment} and \emph{organizational transformation}, as presented in section
\ref{sec:inclusioncriteria}. Preliminary searches had showed that picking
keywords with good precision was difficult. Therefore we decided to exclude the
facet \emph{large scale} from the keyword search, with the consequence of
increasing the manual filtering effort in the subsequent steps.
Having a substantial part of the matches in non-relevant areas such as agile
manufacturing, we included only articles including the term ``software'' or
articles published in relevant conferences. Instead of engineering the search
terms any further, which proved to result in excluding some relevant articles,
we chose to rely more on manual filtering. The resulting facets and keywords are
listed in Table \ref{table:searchterms}.

\begin{table}
    \begin{tabular}{ p{0.22\textwidth} p{0.76\textwidth} }
        \toprule
        Facet                  & Keywords   \\
        \midrule
        Agile methods\newline (before \& after) &
            \texttt{agile, scrum, "extreme programming",
            waterfall, "plan-driven", RUP}\\
        Organizational transformation &
            \texttt{transform*, transiti*, migrat*, journey, adopt*, deploy, introduc*,
            "roll-out", rollout}\\
        Only software\newline related atricles &
            \texttt{(software OR (conference="agile, xp, icgse, icse"))
            AND NOT (title+abs="manufacturing" OR conference="agile manufacturing")}
        \\
        \bottomrule
    \end{tabular}
    \caption{Facets and related search terms used}
    \label{table:searchterms}
\end{table}

The database keyword search matched 1874 unique papers. The abstracts of these
papers were categorized independently by both Dikert and Paasivaara into three
categories: include, excude, and uncertain. The result was combined by marking
abstracts both researchers agreed on for inclusion or exclusion respecitvely.
The inclusion of disagreed or uncertain abstracts was resolved through
discussion. At this stage papers were excluded only if both reserachers deemed
it clearly irrelevant, including any uncertain cases for full text filtering. As
a result 170 papers were selected for full text filtering.

We also evaluated the keyword search against the benchmark created in the
preliminary search step, concluding that 75 of the 117 preliminarily selected
papers matched. The missed 32 preliminary papers were examined, resulting in
including 3 additional papers as primary sources.

% Inter-rater statistics
%
%                     Maria
%            excl   ???   incl    TOT
%      excl  1578    18    15    1611
% Kimi ???     84    17     8     109
%      incl    51    42    62     155
%      TOT   1713    77    85    1875
%
% ???'s =  169
% in/ex = 1706
%
% Pr(Kimi, in) = in / (tot)  =  155 / 1857 = 0.0619
% Pr(Kimi, ex) = ex / (tot)  = 1611 / 1857 = 0.8675
% Pr(Maria, in) = in / (tot) =   85 / 1857 = 0.0458
% Pr(Maria, ex) = ex / (tot) = 1713 / 1857 = 0.9225
%
% Pr(e) = Pr(Kimi, in) * Pr(Maria, in) + Pr(Kimi, ex) * Pr(Maria, ex) = 0.8031
%
% Pr(a) = (ex-ex + in-in) / tot = 0.8831
%
% k = 0.4063

Full text filtering was performed by evaluating each article against the three
facets of the inclusion criteria. Filtering was done in two steps. In the first
step Dikert extracted data relevant to the three facets of the inclusion
criteria. Based on the extracted data 76 papers could be immediately deemed as
included or excluded. The remaining 94 papers were evaluated against the
inclusion criteria by both Dikert and Paasivaara, and a decision was made after
discussing each paper separately. In difficult cases Lassenius was consulted to
reach a decision. As a result of the full text filtering 47 papers were selected
to be included.

In the full text filtering step the references of all papers were also examined
for relevance, including 2 referenced papers as primary sources. Most case study
papers used references very scarcely, mostly referencing well known descriptions
of agile methods.

Finally, counting 47 papers from the full text filtering step, 3 papers from
preliminary searches, and 2 papers from references, 52 papers were selected
as primary studies.

In several cases there were more than one primary study presenting the
transformation of the same organization. Even if the transformation of a single
organization was described in many studies all sources that passed the inclusion
criterial were included. Studies presenting the same organization were treated
as one unit so that we could gain as much insight from each organization as
possible. There were also a few papers that presented multiple case studies, and
in those cases we treated each studied organization individually. As a result 42
unique organizations were identified in the primary studies. We use the term
\emph{study} to refer to the primary study publications, and the term
\emph{case} to refer to an individual case organization that may be described in
several different studies.

% -----------------------

% - pitää olla systemaattinen
% - pitää olla toistettava
% ==> Perustuu enimmäkseen siihen että on 2 tutkijaa toistamassa
% 
% Jos koko on epäselvä:
% --> Tutkijat päättävät keskustelun ja kvalitatiivisen harkinnan perusteella
% --> Tähän tarvitaan esimerkit, milloin on otettu mukaan, milloin on karsittu
%     pois

\subsubsection{Study quality assessment}

The primary research for this literature review consists almost exclusively of
industry experience reports. This may be seen as a factor lessening the
evidence.
--> Why did we do this anyway?
--> Case studies with research method studying organizational transformations in
    software industry are very scarce.
--> Large organizational changes are hard to measure experimentally (controlled
    experiments?). Also due to this any student research was excluded.

The reason for deviating stydy assessment
--> Original research consisted almost solely of experience reports
--> The fact that experience reports have a subjective standpoint makes it
    unreliable to make quantitative conclusions based on the data
--> We also decided to include separate studies describing the same organization.
    This was useful as it deepened the understanding of each case, and several
    large organizations have produced more than one description of their agile
    transformation.

The fact that the study is based on a quantitative basis
--> Experie reports have a subjective bias
--> The data synthesis must not rely on quantitative measures. This means that
    we also considered topics raised by only a marginal number of reports.
--> It may still be interesting to make some quantitative observations such as
    a particular problem being reported in the majority of reports.

It was quite common that a single transformation case was described by many
papers. For example, one paper would highlight the viewpoint of developers
\cite{Fry2007}, and another would consider the transformation from user
experience designers' point of view \cite{Federoff2009}. Instead of using the
most complete paper as suggested by Kitchenham \cite{Kitchenham2007}, we
combined the results presented in each paper and considered the case as a single
unit. By doing this we avoided bias caused by duplicate publications, and were
enabled to have a more detailed view of the case organization.

\subsubsection{Coding of primary documents}

The primary documents were coded using the Atlas TI qualitative data analysis software.
An integrated deductive and inductive approach was chosen for coding the primary
studies \cite{Cruzes2011a}. The coding was designed to have a contextual part
and a findings part as also suggested by Cruzes and Dybå \cite{Cruzes2011a}. A
list of codes was established for contextual information, as presented in Table
\ref{table:contextualcodes}. Codes related to the research questions were chosen
to be created by an inductive process. The reason for inducting codes rather
than using an a priori approach was to avoid the researchers' previous
assumptions of the research area to affect the choice of codes

\begin{table}[h]
    \begin{tabular}{ p{0.28\textwidth} p{0.7\textwidth} }
        \toprule
        Contextual code     & Explanation   \\
        \midrule

Agile method & Agile methods used in the organization (eg. Scrum, XP) \\

Business area & The business area in which the organization operates. \\

Organization size & Mentioning the size of the case organization. \\

Time of\newline transformation & When the transformation has been in
progress, or how long the transformation has taken. Possibly relative to
the time of the research. \\

Research process & The paper describes a research process. \\

Geographical\newline location & Where the organization is located. \\

Large scale definition & A definition of "large scale" software development. \\

Multisite / GSD & Mentioning of a multisite organization. \\

        \bottomrule
    \end{tabular}
    \caption{Context information}
    \label{table:contextualcodes}
\end{table}

To give direction for code induction seven different code families were
established, as presented in Table \ref{table:codefamilies}. A description
defining scope and examples for codes were defined for each family. The example
codes would not necessarily be used as actual codes, and would not represent the
entire or final scope of the families.

\begin{table}[h]
    \begin{tabular}{ p{0.32\textwidth} p{0.65\textwidth} }
        \toprule
        Code family          &  Description
        \\
        \midrule

        RQ1: Reason to change &
        Reasons to start the transformation (demand for faster delivery). \\

		RQ2: Transformation\newline process &
		Statements that describe the transformation process (top-down, big bang,
		stepwise). \\

		RQ3: Challenges &
		Statements that present challenges in the transformation (change resistance).
		\\

		RQ3: Success factors &
		Statements that present success factors in the transformation (management
		support).
		\\

		Investing in change  &
		Factors that present how the organization is investing in the
		transformation (training, consultants, tools). \\

		Practices &
		Distinguishable practices that are used or established during transformation.
		Factors that are presented as neutral in terms of successes and challenges
		in the primary studies.
		(coaching, piloting, continuous integration, communities of practice) \\

		Contextual &
		Contextual codes defined in Table \ref{table:contextualcodes}. \\

        \bottomrule
    \end{tabular}
    \caption{Code families}
    \label{table:codefamilies}
\end{table}

The inductive coding was done according to the guidelines described below.

A passage of text that presents any concept relating to the research questions
or the contextual codes becomes a quotation. If the quotation corresponds
closely to a concept that has already been coded the existing code is used.
Otherwise a new code is created, and the code is assigned to one of the
predefined code families. The labels of existing codes may need to be adjusted
when new quotations are assigned. Many codes may be assigned to one quotation
and quotations of different codes may be overlapping.

The length of the text passage that make up a quotation may vary. For the
contextual codes the quoted passage should be of minimal length. For other
quotes the length of the quotation should be long enough to highlight the
context of the concept the quotation, varying from a few sentences to one
paragraph, or even several paragraphs. As an example of context of the
quotation, if a transformation challenge is described by a quotation it would be
good if also the source or resolution of the challenge could be included in the
quoted text passage. Having enough context should broaden the understanding
of the coded concept when looking at the quotation in isolation..

% -- Considering that mentioning the inclusion of a factor as a success, then
%    would the absence of the same factor be a challenge?

The total accumulated codes and quotations of the integrated deductive and
inductive coding process are presented in Table \ref{table:codecount}. The total
number of quotations is less than the sum of quotations in the categories
because a single quotation may have multiple codes.

\begin{table}[h]
    \begin{tabular}{ l r r }
        \toprule
        Code family    &  Codes  &  Quotations
        \\
        \midrule
        RQ1: Reason to change &        30 &  123 \\
		RQ2: Transformation process &  16 &  580 \\
		RQ3: Challenges &              40 &  323 \\
		RQ3: Success factors &         44 &  260 \\
		Investing in change  &          5 &  137 \\
		Practices &                    11 &  170 \\
		Contextual &                    8 &  215 \\
		Total &                       154 & 1575 \\
        \bottomrule
    \end{tabular}
    \caption{Number of codes and quotations per code family}
    \label{table:codecount}
\end{table}

\subsubsection{Validation of coding}

Coding was done by only one researcher due to resource constraints and volume of
source material. As recommended by Kitchenham \cite{Kitchenham2007} we conducted
an experiment with the goal to assess the consistency of coding that two
independent researchers may achieve.
In the experiment Dikert and Paasivaara coded independently a sample of five
primary documents. The sample was hand picked as a representative subset of the
source material. The coding was done relying solely on the guidelines presented
in the previous chapter, and the researchers were careful not to discuss the
choice of codes or interpretations relating to codes before the experiment was
completed. Finally the independent codings of the two researchers were analyzed
with two metrics: by the number of matching themes found, and by the amount of
overlap in quotations similar codes have.
% See Kitcenham, chapter 6.4

The metric of matching themes measures how well the independent researchers are
able to induct similar codes from the source material. As codes are inducted it
is likely that two researchers label similar quotes differently. To overcome
this problem we defined higher level themes under which the individual codes
were organized. Through discussion we made decisions on whether differently
labeled codes could be organized under the same theme or not. Table
\ref{table:codingexperiment} presents the total number of themes the researchers
found in each document, and how many of the themes were matching. We concluded
that on average two of three themes found by one researcher are similarly found
by the other researcher, but one of three themes is not found by the other
researcher.

The metric of quotation overlap measures how well the quotations for similar
codes match between two independent researchers. The intent is to measure how
codes that are deemed similar by their names match in the actual quoted text. On
a scale of 0 to 3 we evaluated the quoted text of codes in each theme that
matched for both researchers, as described in the previous paragraph. A score of
3 was given for matching codes whose quotation content was virtually the same
for both researchers. A score of 2 was given if the quotations were mostly the
same, but other researcher had included one text passage that the other had
missed. A score of 1 was given if roughly half of the quotations were the same.
If the quotations had less similarity than this a score of 0 was given. One code
provided typically 1 to 3 quotations per document. The quotation overlap score
was counted including only codes that had matching themes. We concluded that the
average quotation overlap score was 2.2.

\begin{table}
    \begin{tabular}{ l r r r }
        \toprule
        Document    &  Matching themes  &  Total themes  &  Quotation overlap (0-3) \\
        \midrule
        P1          &  14  &  19   &  2.1  \\
        P2          &   6  &  11   &  2.4  \\
        P3          &   7  &  11   &  2.1  \\
        P4          &   2  &   4   &  2.3  \\
        P5          &  12  &  16   &  2.0  \\
        \bottomrule
    \end{tabular}
    \caption{Result of independent coding experiment}
    \label{table:codingexperiment}
\end{table}

Combining the quotation overlap result and the result that inductive coding
yielded similar themes with a factor of two thirds, we conclude that two
different researchers should agree completely on at least half of the quotations
in the source material of our study. This agreement includes both that the
researchers select the same text passage as a quotation, and that researchers
agree with the meaning of the quotation. Further interpretation of the
quotations and codes will still remain a subjective matter.

\subsection{Limitations of research method}

-- A keyword search can not identify the facet of \emph{large scale}, and also
   \emph{transformation} is difficult to capture with keywords. Because of this
   we assume that it is likely that a some studies have been missed by the
   keyword search. As an improvement we would suggest combining the entire
   archive of \emph{Agile Conference} and \emph{Agile Processes in Software
   Engineering and Extreme Programming} to the results of the keyword search.
   These two conferences provided more than two thirds of the pimary sources.

-- Few of the primary studies presented a research method. Most of the studies
   were experience reports where the author was a member of the organization
   discussed. Because of this we must assume that the primary studies are
   biased. Due to this risk of bias we chose not to make statistical
   interpretations of the results. We only use relative measure when comparing
   results.

-- We avoid numbers in results, such as frequency of a phenomena, except in the
   meta-analysis section.

% ---------------------------------------------------------------------
\section{Results}
\label{sec:results}

This section presents the findings of the literature review. In the first
subsection a meta analysis of the analyzed papers is presented. In the
subsequent sections findings are presented organized according to the research
questions. The methods used for extracting the results were described in section
\ref{sec:method}. Note that the term \emph{study} is used when refering to a
primary study publication, and the term \emph{case} is used when refering to an
organization which may be described in several studies.

\subsection{Overview of primary studies}

The results of this study include findings from 52 primary studies presenting
how 42 different large software organizations introduce agile methods. Most of
the included studies were experience reports (45 studies), and in many cases it
was evident that the author had been personally involved in the transformation.
Only 6 of the included studies had a research method explicitly stated. One of
the studies was an interview article. The publication forums of the primary
studies were distributed so that 57 sources were conference proceedings, 4
sources were journal articles, and one source was a technical report.

Figure \ref{fig:transformation_time} summarizes the publication years of the
primary studies, and the years when the transformations were reported to have
started. All studies and transformations were dated after year 2000. There were
peaks in transformation studies in years 2008 and 2011. The studies were
typically published two years after the start of the transformation.

\begin{figure}[t]
  \begin{center}
    \includegraphics[width=1\textwidth]{graphics/transformation_time.pdf}
    \caption{Publication years and reported start years of transformations}
    \label{fig:transformation_time}
  \end{center}
\end{figure}

The size of the organizations varied from the minimum included size of 50 to
18,000 people. The median size was 300 people. In 7 studies the size was
presented in terms of teams, ranging from the minimum of 6 teams to 150 teams.
The median was 10 teams. In 8 studies there was no direct indication on size
but the issues discussed revealed that the organization was of large size
accoding to our definition. The distribution of organization sizes is presented
in figure \ref{fig:organization_size}.

\begin{figure}[b]
  \begin{center}
    \includegraphics[width=1\textwidth]{graphics/organization_size.pdf}
    \caption{Distribution of organization sizes in primary studies}
    \label{fig:organization_size}
  \end{center}
\end{figure}

The business areas of the case organizations are presented in table
\ref{table:businessareas}. Different business areas were represented somewhat
evenly. Online services was the largest group, including companies
providing software as a service solutions for businesses, online media players
for consumers, online services for consumers, and communication software for
businesses. The second largest group was telecommunications, including companies
such as British Telecom, Cisco, Ericsson, and Nokia Siemens Networks. The third
largest group was enterprise management solution providers with products for
business process management, project protfolio management, and facility
management.

\begin{table}[!t]
    \begin{tabular}{ l r }
        \toprule
        Business area    &  Case organizations   \\
        \midrule
		Online services for consumers and business  &  9  \\
		Telecommunications                          &  7  \\
		Enterprise management                       &  5  \\
		Banking and financial services              &  4  \\
		Healthcare                                  &  3  \\
		IT services                                 &  2  \\
		Government                                  &  1  \\
		Information security                        &  1  \\
		Not specified                               & 10  \\
        \midrule
		Total                                       & 42  \\
        \bottomrule
    \end{tabular}
    \caption{Business areas of case organizations in primary studies}
    \label{table:businessareas}
\end{table}

\subsection{Agile methods used}

The agile methods reported being used in the transformed organizations are
summarized in table \ref{table:agilemethods}. The most prevalent method was
Scrum, which was the sole agile method mentioned in 25 cases. The second most
mentioned agile method was Extreme Programming (XP). Lean software development
was mentioned in some studies, although combined to Scrum in all cases. Other
agile methods mentioned were Unified Process, ADM, and Rapid Application
Development. In one case the agile method was not named.

It was quite common that organizations sought to combine agile methods.
Especially Scrum, XP, and Lean software development were used together. In
addition, many cases mentioned use of XP practices without explicitly stating XP
as the process being used (such as test driven development and continuous
integration). Combining and customizing agile practices was also evident in that
many organizations viewed the agile method as continuously evolving.
Organizations evolved the agile methods for instance through retrospectives and
continuous improvement.

\begin{table}[t]
    \begin{tabular}{ l r }
        \toprule
        Method                             &  N  \\
        \midrule
        Scrum                              &  25 \\
        XP                                 &  4  \\
        Scrum and XP                       &  5  \\
        Other                              &  4  \\
        Combining Lean to an agile method  &  6  \\
        Not specified                      &  1  \\
        \bottomrule
    \end{tabular}
    \caption{Agile methods reported in number of cases}
    \label{table:agilemethods}
\end{table}

\subsection{Reasons to change}

As stated by Research Question 1 we investigated the reasons organizations start
a transformation process towards agile software development. Most primary
studies mentioned reasons for starting the transformation, and only 7 cases of
42 did not mention any reasons. In the cases giving no specific reason to change
it was simply mentioned that management had decided to change the organization.
The identified reasons for organizations to change are presented in table
\ref{table:reasonstochange}.

% Some cases reported that the organization had been successful in waterfall
% delivery for a long time before the agile transformation.

\begin{table}
    \begin{tabular}{ p{1.2cm} p{1.2cm} l }
        \toprule
        \multicolumn{3}{l}{Reasons to change}  \\
        \midrule

        \multicolumn{3}{l}{External issues: Need to respond to changes in business}  \\

        &  \multicolumn{2}{l}{Demand for more frequent deliveries} \\
        &  &  Time To Market must lessen  \\
        &  &  Long cycle times are seen as problematic  \\

        &  \multicolumn{2}{l}{Something must be done in order to remain competitive} \\
        &  &  Change is required to stay in business  \\

        &  \multicolumn{2}{l}{Customers demand improvement} \\

        &  \multicolumn{2}{l}{Demand for flexibility or responsiveness} \\
        &  &  Demand to accomodate changes efficiently  \\
        &  &  Better responsiveness to customer requests  \\

        \multicolumn{3}{l}{\rule{0pt}{4ex} Internal issues: Problems in managing software development}  \\

        &  \multicolumn{2}{l}{General problems with management of software development} \\
        &  &  Management dysfunction  \\
        &  &  Problems with motivation and morale  \\
        &  &  Problems coordinating a growing organization  \\

        &  \multicolumn{2}{l}{Problems with overhead and excess in process}  \\
        &  &  Excess, slowness and costs caused by process  \\
        &  &  The current process is being ignored  \\

        &  \multicolumn{2}{l}{Problems with planning or keeping schedules}  \\

        &  \multicolumn{2}{l}{Too long cycle time before feedback}  \\

        &  \multicolumn{2}{l}{Demand to improve product quality}  \\

        \multicolumn{3}{l}{\rule{0pt}{4ex} Communication and collaboration issues}  \\

        &  \multicolumn{2}{l}{Communication and collaboration dificulties}  \\
        &  &  Problems in communication across and within projects  \\
        &  &  Teams work in silos or isolation  \\

        &  \multicolumn{2}{l}{Requirements engineering issues}  \\
        &  &  Requirements are not understood  \\
        &  &  Unawareness of what is important  \\

        &  \multicolumn{2}{l}{Demand of better customer collaboration}  \\

        \bottomrule
    \end{tabular}
    \caption{Summarizing why large organizations initiate an agile transformation (RQ1).
             The table presents a hierarchial view of reasons for the organization to change.}
    \label{table:reasonstochange}
\end{table}

The reasons to change were are grouped in three top-level categories: external
issues, internal issues, and communication issues. External issues are related
to the business area which the organization is operating in. Internal issues are
related to the way the organization is managed and software is being developed.
Communication and collaboration issues are related to internal and external
collaboration. Change was motivated by known problems in these categories, such
as known process dysfunctions or communication problems. There was also internal
and external demand for improvement, for instance in speed of delivery and
quality.

\subsubsection{External issues}

The most prominent reason for introducing agile methods was the demand to
deliver software faster and more frequently. A common statement was that it was
necessary to reduce the Time To Market (TTM) [P17, P18, P32, P39, P47]. As
stated in one study: ``The market was moving faster than the development group
was able to deliver'' [P38]. Long cycle time, i.e. time from start of project to
delivery, was seen as a problem in itself [P25, P36]. Some studies indicated
that the organizations set as one of the transformation goals to deliver
software faster [P2, P8]. Other cases described that development was generally
considered as slow [P36, P46, P49].

Commercial competition and market situation was seen as a trigger for
transformation. Especially speed of delivery was seen as a means to respond to
the competition [P9, P38]. As Greening describes: ``The average release duration
peaked at 41 and 35 months, an alarming state that could enable competitors to
gain market share.'' [P19]. In many cases the transformation was seen as
essential for the company to remain competitive [P10, P11, P18, P42, P49]. For
instance, British Telecom realized that ``From a strategic view, there is little
alternative than to change the old ways of working'' [P32].

Demands for change and improvement were also expressed by internal and external
customers. Again, the requests of the customers reflected the need of faster
development and release cycles [P5, P39]. As described in one study: ``Speed of
delivery is of the essence – customers are no longer prepared to wait for 18
months or more for solutions to be delivered'' [P23]. In some cases business
partners expressed mistrust in the original way of conducting development [P40,
P46].

Yet another frequent external factor driving for improvement was the need to
accomodate change during development. Organizations saw agile as a opportunity
to improve responsiveness to change in general [P25, P26, P42].
Some organizations faced difficulties when changes to plans were needed. For
instance, the organizational struture made adapting to changes difficult [52],
and making changes created management overhead [P51]. Other organizations saw an
business advantage if priorities could be changed as required [P8, P39].
Customers and business representatives were in many cases attributed as the
sources of change requests, and the responsiveness to customer requests needed
to be improved [P18, P33, P36, P49].

\subsubsection{Internal issues}

The internal issues relate to how software development is managed in the
organization. In many organizations the project management practice was in a
state of dysfunction, and it needed to be corrected. Management problems were
for instance constantly recurring crisis events [P30], managers ignoring
prioritization of tasks [P38], and reluctance towards focusing on delivering
business value [P49]. Some cases stated that projects simply failed [P21]. Some
organizations wanted to improve generic project management practice, such as
risk management [P8], management of offshoring [P38], and improving the quality
of life of team members [P44]. Several organizations were growing and
experienced that the old development method did not scale up [P16, P30, P50, P51].
There was also a desire to improve team motivation and commitment, and
collaboration between management and developers [P39].

Command and control style management where ``a few people, usually those
farthest from the action, made the decisions without directly involving the
developers doing the actual work'' was highlighted as a problem [P44]. For
instance, schedules could be dictated by management, and teams had to comply to
them regardless of the workload [P9]. As a result teams were overworked and
needed stretch far to finish projects on time. Low morale and motivation were
identified as problems that needed to be fixed, and agile was seen as a solution
[P25, P30, P40].

A common problem that was acknowledged in organizations was overhead in process,
which showed as bureucracy and needless costs. Examples on overhead in process
were unproductive meetings [P38], sign-offs and gates [P9], excessive focus on
conformance to process [P40], and change management overhead [P51]. Creating
excessive documentation instead of business value was seen as a problem in some
cases [P24, P35].
From business representatives point of view the processes were unreasonably
costly, while at the same time insufficiently capable of delivering business
value [P36, P46, P49]. In one case business realized that customers did not
request the heavy CMMI based process [P24].
A heavy process resulted often in ignoring the process, only paying ``lip
service'' to the process, or applying its steps retroactively [P6, P9, P49].
Seffernick described how applying the waterfall process had teams ``still
developing the same old way but slower because of the additional process''
[P46].

A slow process with long cycle times leading to late feedback was a problem in
the old model. A large investment was required before results were visible [P5],
and there was variances in expectations and what was delivered [P40]. A related
problem was late integration which caused extra work [P25]. As organizations
grew and products became more complex test cycles began to lengthen, which
hindered flexibility and reduced speed [P33, P50].

Yet another problem in managing development was problems with schedules.
Some cases described schedules as unpredictable and estimation being difficult
[P16, P30, P40, P51]. In other cases it was reported that release deadlines were
often missed [P19, P21, P33, P41].

Finally, the original process model was seen as a cause for quality problems.
Several studies mentioned a demand for better quality, and quality improvement
was set as a goal for the agile transformation [P25, P31, P41, P42]. Improvement
in external quality and customer satisfaction were also anticipated from agile
development [P2, P4]. Problems with technical debt were attributed to the
original process. Some cases reported on prioritizing new feature development
over bugfixes or fixing only the easiest issues [P30, P38].

\subsubsection{Communication and collaboration issues}

Many organizations experienced problems in communication and collaboration
between projects, teams, and even members of a single team [P38, P21].
Teams were in many cases organized around components and were working in silos.
Working in silos diluted focus and made adapting to change difficult, as teams
could only focus on their limited area of expertise [P7]. Functional silos acted
against creating a shared purpose in operating [P40], and resulted in a ``ball
over the wall'' mentality [P46]. Distance between teams was also created by
document driven communication, and was causing misunderstanding of requirements
[P24].

Other communication issues were related to requirements handling and difficulty
to focus on the most important aspects of a delivery.
Examples on problems with requirements was too much focus on the documenting
aspect instead of providing understanding [P24, P49], and ignoring the fact that
requirements and understanding are evolving during a project [P4].
In some cases the organization of work and management was such that it hindered
focusing the development [P7], and prioritization of what to work on was not
done efficiently [P38, P39].

Too little customer and user collaboration led to problems in some
organizations. For instance, too late user involvment led to change requests
late in the project [P40]. There was a wish to collaborate more frequently and
in less formal ways [P4, P26].

A few organizations were experiencing rise of agile methods at grass roots
level, but the agile teams could not collaborate efficiently with other parts of
a waterfall organization. Tensions arose when teams using different development
approaches attempted to work together on products [P10]. On a team level agile
had enabled some flexibility, but that flexibility could not be leveraged on the
scale of the entire organization [P26]. This kind of experiences drove the
organization to deploy agile on a full scale.

\subsection{Transformation process}

\subsubsection{The initial state of the organization}

We investigated the initial state of the case organizations before the
transformation. The typical initial process model was described as waterfall, as
stated in 25 cases. In several cases the authors characterized the initial state
of the organization as ``traditional''. CMMI was presented as a driver for the
original process in 4 cases. Other process models the studies named were RUP
[P19], Prince2 [P51], and Macroscope [P14]. In 9 cases the initial state was not
described.

The old model was described in most cases briefly and in a neutral sentiment.
Waterfall development in general was referred to as ``traditional, plan-based
methods'' [P25], ``traditional phase-based approach'' [P36], or ``old fashioned
process driven waterfall'' [P23]. Korhonen [P25] referred to the model presented
by Royce \cite{Royce1970} when describing the initial state.

In some cases specifics of the old model was criticized. The characteristics of
the old process models were described as having ``many gates and sign-offs
required by upper levels of management'' [P9], ``lot of bureaucracy in project
management and a strict hierarchy of requirements communication'' [P39], and
``project managers typically used a command-and-control philosophy'' [P44].

In a few cases the initial process model was less rigid, and closer to the agile
model that was to be introduced. Examples of such loose process models were
described: ``The choice of development methodology is one of any number of
degrees of freedom granted to teams at Amazon. Development teams are expected to
solve their problems in the best way possible'' [P3], ``the R\&D group leveraged
a loose, waterfall-based process with an entrepreneurial culture'' [P16]. Some
organizations had even signs of agile development on team level although the
surrounding organization operated in a waterfall fashion. This was described as
``Agile and Scrum were increasingly popular and emerging at a grass-roots level,
at a corporate level Yahoo! still embraced a more traditional product
development process'' [P10], and ``some teams had applied scrum on the team
level, despite the surrounding organization that was working with a waterfall
software engineering model'' [P26].

In a few cases the organization initailly lacked strict process. For instance
Long describes the initial state in the following way: ``Project scope was in
constant flux. Well-meaning market and product owners routinely inserted new
requirements during product development.'' [P30].

Some authors describe the developement being originally organized in silos. For
instance Ranganath describes that ``Teams operated in silos. Collaboration,
shared purpose and ownership for outcomes were hard to find'' [P40].

\subsubsection{Modes of transformation}

Work in progress...

Top down

Bottom up

Big bang

Transitional

\subsubsection{Typical activities in the transformation}

Training

Use of external consultants

Coaching

Pilot program

Seminars, workshops, other public events

Agile community

\clearpage

\subsection{Challenges in transformation}

Any organizational transformation that involves numerous individuals will face
some challenges. As stated by Research Question 3, we investigated what
challenges large organizations introducing agile methods may face.

The challenges presented in this section are factors that primary studies
described hindering the progress of the transformation. Some of the challenges
presented here do not explicitly relate to the transformation process, but are
rather general difficulties in agile imlementation. These issues are still
included as settling the agile implementation can be seen as a part of the
transformation process. Some solutions for challenges are also presented, but
they should be considered as ways of describing the challenges rather than
direct success factors.

\begin{figure}[!b]
  \begin{center}
    \includegraphics{graphics/challenges_summary.pdf}
    \caption{Overview of challenges in agile transformation}
    \label{fig:challenges_summary}
  \end{center}
\end{figure}

The challenges found in the primary studies are summarized in figure
\ref{fig:challenges_summary}. The challenges are organized on a high level into
seven categories within three themes. The categories are discussed subsequently
in this section. First we present generic transformation issues that may apply
to any organization. As the second theme challenges relating particularly to
sizable organizations are discussed. The final theme studied is challenges
encountered when agile development needs to interface with other functions
involved in a system development life cycle.

It is to be noted that we do not emphasize any of the listed categories or
challenges above another. The challenge categories we present were found evenly
in the set of primary studies. We included only challenges that had been
mentioned in several studies, or were particularly emphasized in fewer studies.
Challenges with only sporadic mentionings were considered less significant and
were left out. There was no further evidence that any challenge would be more
significant than another.

\subsubsection{Change resistance}

In an organizational transformation working habits and roles need to change.
However, people are usually disinclined to change and some resistance should be
anticipated in any change process. People will not be willing to change unless
there are good reasons, and the change is preceived easy enough. It is difficult
to attain a buy-in for a change, and even organizations that have a flexible
culture will face resistance [P10]. It should be expected that not everyone will
be willing to change, even so that some employees will never adapt to the new
model [P6]. During a transformation period objections to change may become a
major reason for loss in time and productivity [P29]. The main types of change
resistance discovered in the primary studies are summarized in table
\ref{table:challenges_changeresistance}.

\begin{table}[b]
    \begin{tabular}{ l l }
        \toprule
        Types of change resistance           &  Described in  \\
        \midrule
        General resistance to change         &  P14, P18, P22, P44, P45   \\
        Skepticism towards a new model       &  P5, P8, P9, P18, P36, P46 \\
        Top down mandate creates resistance  &  P2, P12, P37, P49 \\
        Management not willing to change     &  P2, P3, P6, P49 \\
        \bottomrule
    \end{tabular}
    \caption{Types of change resistance reported in studies}
    \label{table:challenges_changeresistance}
\end{table}

Numerous reasons were given for why change resistance emerged. For
instance, Fecarotta [P14] described the organization of Boeing as risk-averse
and cautious, which slows down any change. Any change was further hindered by a
deeply rooted status-quo and high employee retention [P14]. In some cases it was
reported that people worried about new roles and responsibilities that the agile
model might bring [P44]. For instance, testers were worried that they would need
to take on cross functional tasks, which would be outside their area of
expertise [P18]. Yet another reason for change resistance was the move from own
offices to team areas [P22]. People were also feeling being monitored because of
the increased level of interaction within the team and between project
stakeholders [P45].

In these cases change resistance was mitigated for instance by education [P45]
and one-to-one discussions [P22]. Change resistance was also softened by first
engaging only teams willing to try agile, so that insights could be gathered to
ease the continued transformation [P45].

Another common form of change resistance that was reported was skepticism and
distrust in suitability of agile. Seffernick describes how management did on one hand
acknowledge benefits of agile, but on the other explained away the possibilitiy
to apply it due to contract reasons, the current matrix organization, and other
organizational practices [P46]. Skepticism was often created by other similar
misconceptions, including that agile does not work for complex products [P5,
P36], agile needs to be implemented in a prescriptive by-the-book way [P9], that
frequent meetings will cause overhead [P18], and that agile equals avoiding
governance and working without a plan [P8].

% For instance, teams may not be willing to try
% new approaches unless they have been proven [P6].

% Resistance is to be expected --> there are many different stakeholders [P41]

% * No value seen by some groups, e.g. mainframe or maintenance [P49]

The way the transformation is initiated will affect how change resistance will
show. In several cases the change initiative came from management, but when
presented wrong the initiative becomes a mandate that pepole are not receptive
for. Spayd [P49] summarizes this as: ``Organizations do not change merely
because the boss says so, at least not in the way that is intended''. In that
organization the introduction of CMM had been carried out consistently by a
mandate, but the collaborative nature of agile did not mix well with such a
mindset. A top down mandate may also dilute the understanding of the reasons for
starting the transformation, and the understanding of agile development overall
[P2]. O'Connor describes how team members became sceptical towards agile when
implementing mandaetd changes did not bring any visible benefits [P37].
A problem relating to top down mandate is that if management does not speak out
a goal towards using agile, developers may feel that the agile methods may be
replaced by something else at any time [P12].

The direction in which a tension between management and development is built may
also be opposite. In cases where the change was emerging bottom up the
reluctancy to change was not in the team members but management, causing that
significant organizational change could not be raised above the team level. For
instance, Atlas [P3] describes how executive support would have been required to
extend the agile process to involve the product management office and separate
quality assurance groups. In this case, when managers %for multi team projects
were not involved in the transformation, the agile model was restrained from
spreading beyond the level of development teams [P3]. Lack of middle management
support for change and disinclination to change management culture were seen as
some of the most serious problems in the transformation [P2, P49]. Middle
management is in a position to undermine the entire transformation, and may do
so if they do not participate in, and thus understand, the agile method. Agile
brings changes to some management roles [P2], and lack of understanding of agile
development will leave managers feeling left out [P6].

\subsubsection{Lack of investment in transformation}

It is clear that changing the way of working will create some costs for the
organization. Problems will ensue if there is no intention to allow resources
for the transformation. Table \ref{table:challenges_lackofinvestment} presents
investments and resourcing arrangements that were witheld, and which caused
problems in transformation.

\begin{table}[b]
    \begin{tabular}{ l l }
        \toprule
        Factors of investment and resourcing  &  Described in \\
        \midrule
        Lack of coaching                      &  P1, P6, P23, P45, P47, P49  \\
        Lack of training                      &  P11, P20, P45  \\
        Excessive commitment of people        &  P37, P42, P49  \\
        Old commitments are kept              &  P11, P21  \\
        Physical spaces can not be rearranged  &  P6, P29, P36, P49  \\
        \bottomrule
    \end{tabular}
    \caption{Summary of investment and resourcing factors that affect
             transformation negatively}
    \label{table:challenges_lackofinvestment}
\end{table}

% Lack of training and coaching

Training and coaching are direct investments in transformation and their lack is
an evident problem. These activities were still underbudgeted, which
created difficulties in the transformation [P11].
Hajjdiab [P20] describes how reluctance of management to invest in training
caused teams to be ill prepared for the agile transformation effort, which
eventually resulted in ending the use of agile methods. As a less dramatic
outcome of too little training was lower motivation [P45].

Arranging proper coaching was a particular problem in many organizations. It is
critical to coach teams in their real work environment as a proper change in
mindset is difficult to achieve by only attending training sessions. [P9, P23]
In large
organizations where numerous teams needed to be coached the demand often
exceeded the capacity of available coaches [P1, P6, P23, P49]. Lack of coaching
was also attributed as a reason why the success of pilot teams could not be
repeated when agile was adopted more widely [P9, P45].
Silva and Doss [P47] described that when it was hard to fill the coaching
positions people less seasoned in agile were appointed as coaches, which risked
that agile practices would not be taught correctly.

% Add quote?
% * No coaching --> Applying techniques improperly --> Damage
% [P9]
% Without proper coaching however, the success of these informal teams met with
% mixed results over the longer term.  In retrospect, many of these teams did
% damage to themselves by improperly applying Agile techniques, particularly when
% command and control mechanisms were still in place.

% Old work and commitments, schedule pressure

Even though the organization was in a state of transformation the workload of
teams was not always adjusted to facilitate the process.
Some organizations started their agile journey in a state where people were
overcommitted. It was later realized that when people are overloaded they will
not be able to changing their behavior and learn new ways of working [P37, P42].
Spayd [P49] also describes how senior management pressed teams to deliver what
the customer had been promised, regardless of what the current velocity of
development predicted. For one part the pressure to deliver interfered with the
transformation, but it was also forcing a change in the old way of working
[P49].

In one case management forced people to remain committed to firm deadlines, even
midst in the organizational transformation [P21]. The engagement in old
commitments and tasks resulted in ignoring new practices introduced by agile,
and eventually the agile method was breaking down [P21].
Cowan [P11] describes how time was allocated for tending to old commitments.
However, even when there was preparation for extra work people with specific
expertise became overloaded. After time it was realized that overcommitment must
be reduced by allowing to shift excess work for later time [P11].

% Physical spaces, challenge of finding a seating

In some cases the old way of working had people seated physically distributed,
in opposite to what agile methods typically suggest. Office spaces needed to be
arranged so that the entire team could work in a single room, but it took time
and effort, and it was not always possible [P36, P6]. Lewis and Neher describe
how dedicated rooms for teams could not be arranged, and team events such as
daily standups required booking a conference room [P29]. In another case the
estate organization shut down the teams' attempts to modify working spaces
[P49].

\subsubsection{Agile is found to be hard}

A challenge most organizations came across was that implementing the agile model
was found to be difficult. An experienced software team may do well in training,
but when the time comes to apply agile techniques in practice the team may get
lost [P21]. Viewpoints on the difficulty of correctly implementing an agile
model are summarized in table \ref{table:challenges_difficulty}.

% The difficulty in implementing agile is evident in misconceptions of the agile
% model, when by-the-book implementations prove to be insufficient, and when the
% agile methods were applied badly to the organization.

\begin{table}
    \begin{tabular}{ p{0.61\textwidth} p{0.31\textwidth} }
        \toprule
        Type of difficulty in implementation  &  Described in \\
        \midrule
        Misunderstanding agile development concepts  &
        P6, P26, P37, P38, P42, \newline
        P44, P45, P48, P49      \\
        Literature does not provide enough guidance  &
        P5, P6, P10, P13, P21, \newline
        P27, P30, P45, P49     \\
        Agile practices are customized badly   &
        P9, P29, P44, P46  \\
        \bottomrule
    \end{tabular}
    \caption{Summary of viewpoints why agile implementation is difficult}
    \label{table:challenges_difficulty}
\end{table}

% Misconceptions of agile

There were many examples on problems caused by misconceptions of what agile
software development is. On a general level, Bang [P4] describes how the values
of the agile manifesto were not understood, and agile practices were carried out
without understanding their purpose. In one case management saw the purpose of
agile simply being faster product delivery [P9].

On a team level there were several examples of misunderstanding agile
development. Cases describe how agile was seen as freedom to hack without
documentation [P26], that development could be done without design tasks [P37],
and that agile meant that everyone should become a generalist [P42]. Schatz and
Abdelshafi describe how teams presented unfinished work at reviews and ignored
the principle of showing only completed increments, which resulted in a backlog
of bugs [P44]. In one case team members had perceived the introduction of agile
as a means to squeeze out more efficiency [P45]. In addition, the flatter
organization was seen as fewer career opportunities, and pushed team members to
internal competition for visibility [P45].
Still one misunderstanding of agile was due to viewing it only through the tools
used, such as project management software [P48]. Superficial focusing only on
tools themselves and not the reasons behind their use led to frustration [P42].

In some cases the misunderstanding of the agile model was described as ``doing
mini waterfalls'' [P6]. Managers were using Scrum terminology, but at the same
time committing unreasonable workloads on teams [P6]. In another case a
waterfall nature was evident when task estimates were given as hours left, and
task breakdowns disconnected from what was really being done [P27].

As a final example of misconceptions, some organizations considered agile being
a solution to all problems [P9]. Success was declared prematurely [P49] but
expectations created by successful early experiments were not fullfilled [P38].
Disapointment was granted when these hopes were not fullfilled, and instead
implementing the agile model was found to be hard and time consuming.

% By the book is not good

Even if there was no misunderstandings the correct agile model was hard to
implement. Several cases describe how an agile model was hard to learn from
literature [P21, P27], especially when it comes to implementing it on a large
scale [P13]. As Beavers [P5] writes: ``There simply was not a manual or document
where we could find easy answers on how to do things''.
Schnitter and Mackert [P45] report that theoretical considerations on how to
scale up the agile methos were not good enough, and that product managers and
architects struggled when several Scrum teams were working concurrently.
Another case concluded that all practices suggested by XP did not fit enterprise
scale development, and some practices required customization [P49].

% In a large organization with several different teams and projects it is clear
% that one single model can not fit all needs [P41]. The model of
% agile should not be prescribed in a strict way, as it would impose a single
% model in an organization with several individual teams and projects [P41].

% Check this paragraph
Further indications were given that the advice from literature will not ferry
all the way. Benefield [P6] describes how it was dificult finding a balance
between prescribing a by-the-book implementation which may put people off, and
giving too much freedom in the agile methods which may weaken core practices.
The reality where the practices needed to be applied was described as messy in
comparison to the ideal circumstances presented in literature [P10]. Because of
this it was difficult to choose a model to start with and gain buy-in for [P10].
Difficulties in choosing a initial model were also due to variances perceived in
the agile approaches suggested in literature [P30].

% Where would this belong?
% Organizations had also bad experiences if the implementation of the agile model
% was being prescribed. For instance, implementing Scrum based on checklists of
% practices would not work [P41]. Benefield described that it was challenging to
% find a balance between independence for teams to choose their way and diluting
% the value and structure of Scrum [P6]. Giving room for team autonomy is one of
% the critical facets of agile, but it must be ensured that teams do not discard
% practices that add value and structure.

% Customizing badly (mapping to old model)

The difficulty of agile and misunderstandings were also evident in cases where
the methods were customized badly. As by-the-book implementation often was not
feasible, the agile method was tailored to suit the needs of the organization.
However, in some cases tailoring meant skipping some practices, which lead into
problems. In one case certain individuals were allowed to ignore some core
elements of Scrum, which came to turn the teams decision making into a variant
of command and control [P9].
Lewis and Neher [P29] describe how there was temptation to strip some practices
and enhance others. Previous attempts had proven that one of the reasons for
agile implementations to fail was deviations from the process, because of which
the agile mindset did not take root [P29]. A bad customization may lead to the
team taking only practices that reflect their current needs, and therefore
failing to achieve a real change in process and mindset [P9, P46].
Another example of bad customization that counteracts change is replacing agile
vocabulary in favor of familiar terminology from the old model [P44]. An
important benefit of introducing new vocabulary is that it stimulates new ways
of thinking [P44].

\subsubsection{Coordinating work between many teams}

In large software projects work must inevitably be distributed between many
teams. One of the most prominent transformation challenges was the difficulty in
creating a model to coordinate work between many teams. Table
\ref{table:challenges_coordinatingteams} presents challenges that arose when
agile was implemented in environments where many teams need to collaborate.

\begin{table}
    \begin{tabular}{ p{0.61\textwidth} p{0.31\textwidth} }
        \toprule
        Problems in an environment with many teams  &  Described in \\
        \midrule
        Interfacing between teams is difficult  & 
        P9, P10, P13, P17, \newline
        P24, P26 \\
        A model with autonomous teams is not good  &  P7, P13, P29, P51  \\
        Distributed setting causes additional problems  &  P29, P45, P48  \\
        Technical integration and consistency  &  P5, P24, P26, P27, P50  \\
        \bottomrule
    \end{tabular}
    \caption{Summary of challenges in a setting with many agile teams}
    \label{table:challenges_coordinatingteams}
\end{table}

Many cases reported that introducing agile practices soon yielded improvements
on the team level, but as a part of a large transformation teams also needed to
change their external behavior.
The initial success was followed by challenges when teams needed to work with
other teams and as parts of the larger surrounding orgaization [P13, P17, P24].
Introducing agile had created flexibility on a team level, but it could not be
fully utilized as the surrounding organization was not responsive enough [P26].
The rollout of agile had not removed dependencies, and the dependencies
made managing development difficult [P9, P10].

% Independence of teams

Some organizations attempted to initially create a model where teams operate
autonomously, which is well in line with agile principles. However, a number of
issues arose from independence. For instance, teams needed to find a balance
between the their own goals and broader goals of the entire organization, but
some teams chose to focus only on their own goal [P7]. Coordination was
obstructed by independent teams that did not respect the larger context [P13].
Coordination was also difficult when teams were working independently for
different customers but the applications being built were interdependent [P29].
One case reported that even allowing teams to have different sprint durations
created delays in delivery [P51].

Further coordination problems were encountered when scaling up agile over many
sites. Operating in a distributed environment makes communication more
difficult, and causes problems when frequent and effective communication needs
to be achieved. Distribution had negative effects, such as missing kick-off
meetings, reduced feeling of proximity when telecommunication is necessary, and
difficulty in arranging frequent meetings due to time zone differences [P45].
Agile project management was also problematic in a distributed setting. In a
waterfall model separate parts of projects could be isolated to different sites,
but the agile model does not allow such strict slicing of projects [P29]. In one
case it was simply admitted that a distributed organization will impose
additional burden on communication and require additional care in the process,
but distribution and agile would still be used together [P48].

Also technical problems relating to inter-team coordination were reported.
Integrating the products of teams was seen as a problem in some cases [P5, P50],
and lack of standardized build scripts was mentioned as a problem in one case
[P27]. There was also problems in synchronizing definitions of software
interfaces between teams, and dependencies in code caused problems in larger
features [P26]. One case reported that team centricity in the agile model
created a fragile architecture, divergence in coding style, and even distrust
between teams [P24].

Moore and Spens [P34] describe the progression of the above mentioned
coordination problems. At first cross-team dependencies were attempted to be
reduced, but it became evident that the dependencies were an inherent part of
the project. A traditional approach to managing dependencies created excess
work, reduced independence and flexibility of teams, and created contract based
and adversarial relationships within the organization. There was some success
when up to 5 teams were coordinated using Scrum-of-Scrums, but that model could
not be scaled to a global level. The main problem was teams that were thinking
too much inside the team walls. When the new practices were scaled up the
communication channels narrowed and created communication breakdowns. Problems
with collaboration, dependencies, and integration were solved in the presnce of
individuals having a proactive and open mindset towards working over team
boundaries. It was concluded that a balance is needed between completing new
stories from the team backlog and maintaining overall stability of the
application. [P34]

Many studies presented solutions for the inter-team challenges, which also
highlight the nature of the challenges themselves. The general model of solving
inter-team problems was introducing an additional layer of coordination, such as
a Scrum-of-Scrums team [P43], a similar product team [P45], or an entire
customized release framework [P13]. In one case integration problems were
controlled by prioritizing code quality and maintaining existing functionality
over new features [P13]. Finally, the choice of tools was reported to have an
impact on the coordination effort [P17].

\subsubsection{Large organization issues}

The primary studies revealed challenges that relate particularly to large
scale. Additional layers of management are usually necessary to steer large
groups of people, but agile methods provide little guidance on how to organize
management. Problems may arise in a large environment if the schedule of agile
adoption varies from team to team. Internal fragmenttation is a further problem
that may be encountered in large organizations.

% Roles of managers

The existence of additional management positions may pose problems to agile
processes that emphasize self organization.
Especially the role of middle management was seen as unclear in agile methods
[P2]. This is problematic as an agile transformation requires cultural change
particularly on the middle management level [P2].
It was described that managers needed to resist the tendency to take command and
control and allow room for self organization, but the change in role was
difficult to achieve for the people involved [P11, P52].
Nielsen and McMunn [P30] describe how the project management group had
previously worked through big up front plans and competing of resources, but
those ways would need to dramatically change.

Several other problems related to management roles were also presented. For
instance, Spayd [P49] describes how managers were left outside the roles offered
by the new agile model. In another case when managers were appointed as Scrum
masters developers felt being micromanaged [P27]. This was partially because of
vague understanding of the agile method [P27]. It was also described how mixing
the role of project manager and Scrum master created a conflict in interest, and
turned the Scrum manager's role into one of policing the team instead of
supporting the team [P30]. In some cases problems with manager positions were
solved by phasing out roles relating to the old model, and replacing them with
new positions more suitable for agile [P38, P52].

% Using two models side by side
% Part of the organization is in old mindset

In most cases a large organizational transformation proceeded gradually, and
during the process it was possible that the new agile model was used in parallel
with the old methods. The arrangement of using the old and new models side by
side was seen as generally problematic [P1, P26], and causing tensions on all
organizational levels [P10]. Ongoing projects had been set up with previous
development methods, and the agile methods needed to be arranged to fit along
with them [P8, P29]. A particular problem in collaborating between projects
using different models was that in Scrum the software design was fleshed out
over time as sprints progressed, but the waterfall model would require detailed
design to be locked early in the project [P29].

Even when agile had been taken into use in development there were problems with
management working according to the old waterfall model. O'Connor [P38]
described management as ``focused on meetings and big up front project plans''
despite agile being in use. In another case management was losing confidence in
agile because familiar reports on cost analysis and percentage complete were not
produced [P35]. As Scrum teams did not commit to fixed schedules they were
considered unreliable according to the original model [P45].

Some cases reported that development efforts were still controlled on a top
level by a project management office (PMO). The PMO was described as entrenched
in a waterfall paradigm, and the top-level rigidity was causing friction in the
agile adoption [P3, P6]. The PMO was seen as a hub and bottleneck controlling
all aspects of projects, and its central structure had to be dismantled for the
organization to become agile [P46].

There were also problems with duplicating management when two models were in
effect. For instance, agile teams were required to comply with current
procedures producing excess documentation and stepping through approval gates
[P20, P21]. The bureucracy of the previous traditional model was still enforced,
and management was not convinced to lighten the process. Another case describes
that after introducing the agile method teams were required to fill in templates
of two processes [P48].

% Challenge: Differing vocabularies

In addition to teams using agile and waterfall models side by side a large
organization makes it possible that the agile interpretations of teams diverge.
When many teams implement agile without consistent guidance friction and
fragmentation may emerge [P9]. The organization may require moving people
between teams from time to time, and therefore it would be desirable that no big
differences in teams' agile cultures exist. Divergence in process would create
risk of increased costs when relocating people [P38].
Further, if processes of teams diverged management had difficulties in
forecasting and comparing work of different teams [P38, P48]. To overcome
problems with divergence in agile models some organizations defined standards to
follow [P8, P43].

% Challenge: Previous organization in silos is haunting new model
% Is this good??
In some cases the initial organization had internal boundaries or specialized
knowledge was stuck silos. Such settings caused problems in the agile
implementation.
Cloke [P10] describes how use of Scrum revealed an internal segmentation where
teams operated with differing priorities and agendas. This segmentation also
hampered the initial effort to use Scrum in a larger context [P10].
Cowan [P11] describes how projects needed specialized skills that were scarce
and people were often relocated to match the needs of the moment. Too much
reorganization made it diffucult for teams to plan ahead [P11].

%% This??
% A dispersed matrix organization causes problems when people are on one hand
% kept assigned to many projects simultaneously, but on the other hand functions
% such as database administration, architecture boards, and project management are
% centralized [P14].

%% ?? -->
% Large organization put agile methods under the test of scaling up, and problems
% were encountered.
% In the agile model it was hard for managers to get an overview of the progress
% of the release [P17]. The status of features was hard to tell because different
% aspects of features were sepatated across teams [P17].
%
% One case reported use of agile across hundreds of teams, but management could
% not get a roll up of the progress without having strict rules on how status was
% reported [P48]. Schnitter and Mackert [P45] report that a maximum workable size
% in a project managed by Scrum was 130 people including 7 development teams.
% Smits and Rilliet [P48] report that providing executive overview of the
% development programs was limited by the tools used to manage agile projects.

% --

% Add quote ?
% * Especially higher organizational levels were not using agile --> Problem
% [P3]
% Scrum still exists mostly in a way that avoids conflict with the larger
% corporate organization, and there has been little progress in addressing larger,
% organizational impediments.

% [P18]
% Real Agile teams will experience resistance when they start working with
% non-Agile interfaces. Lack of understanding in ‘why to help create a cadence?’,
% ‘why involve the customer more?’ will cause resistance in the change process.

% --> ``Part is in the old mindset''
% * Problem: Management attempts to use a ``resource pool'' of individuals rather than coordinating teams
% [P6]
% we face the constant threat that restructuring and the instinct from management
% to track and leverage resources at the individual level (for example
% organization-wide resource pools) will destroy the team dynamic and adversely
% affect productivity.

\subsubsection{Integrating other functions than development to agile}

Agile methods are typically considered as a set of activities closely related to
software development. Cases describe how development teams were the first
entities in the organization that were considered when introducing agile [P2,
P17, P31, P38]. However, development operates in a larger context and will need
to interface with other organizational functions.
Challenges were faced when exposing other parts of the organization to the agile
model, and functions that were further away from development were resistant to
change [P2, P18, P31]. A tension was growing between development and broader
roles in the organization [P1]. The full beenfits of the transformation would
not be attained unless the entire organization was set to work along the same
paradigm [P18, P42].

Primary sources listed various functions beyond development that are involved in
bringing products to the market. Functions interfacing with development and
affected by the agile transformation are presented in figure
\ref{fig:challenges_other_functions}.
Marketing was a particularly frequently mentioned group [P5, P17, P28, P31, P38,
P45, P48]. Other groups involved in the whole of product delivery were sales
[P17, P31], infrastructure and operations [P10, P45, P46, P50], user experience
and design [P9, P10, P28], documentation [P28], legal [P10], security [P10],
customer services [P48], and finance [P48].

\begin{figure}
  \begin{center}
    \includegraphics{graphics/challenges_other_functions.pdf}
    \caption{Functions affected by the agile transformation}
    \label{fig:challenges_other_functions}
  \end{center}
\end{figure}

Agile affects timings in the software delivery lifecycle which caused two types
of problems. Firstly, the iterativeness of agile models was problematic when
interfacing with other functions. Iterativeness changed the pace of delivery
[P17], and forced design to focus on a shorter scope [P15]. Various groups
feeding and supporting development had earlier committed once to a long term
plan, but in the new arrangement they needed to adjust to incremental delivery
[P48].
Especially user experience (UX) and interaction designers did not welcome an
incremental model [P6], and were struggling with maintaining the big picture in
design while working in iterations [P9, P10].
The infrastructure and operations teams were falling behind due to the increased
delivery speed of development teams, and they were forced to update their way of
working [P46, P50].

Federoff and Courage [P15] elaborate on the user experience team's problems
with the short time horizons of Scrum. Previously the UX team had supplied
development with designs within a long release cycle, but Scrum required
completing features within weeks. The problem was that that the UX work required
a holistic view instead of an incremental one, and the design process did not
fit well into the time frame of a sprint. The UX team was overloaded and
discontent, but the situation was fixed by refining the schedule for interations
with development teams. [P15, P28]

The second type of problems was that some release activities would inevitably
have a long lead time but required committing to a set of functionalities in
good time. The agile models' emphasis on short time horizon and flexibility in
prioritization did not mix well with such activities.
For instance, it would take time to prepare printed marketing material and press
releases, but they needed to accurately present the upcoming features of the
product [P38].

Maples [P31] describes that marketing needed three months for preparing the
product launch, but the agile process allowed development to keep changing the
content of the product. As a result marketing was struggling with creating
material and campaigns without knowing the exact product features. Further, the
processes of acquiring export clearances and licensing authority required that
the the feature set of the product was known well on beforehand.
As a result of these requests development felt that their process was being
slowed down. The issues were solved by agreeing on that marketing could at any
time request a release in 90 days, and development would be able to commit to a
fixed feature set to be delivered [P31].

% Problems with HR

Problems in adopting agile were even showing in functions outside the software
delivery lifecycle. In several cases human resouces (HR) was a support function
mentioned to work against the agile adoption. In order to gain full results of
the transformation the entire organization should to be aligned, including HR
[P38, P42, P49].
Especially the rewarding practices set by HR were seen as problematic for the
agile model. Rewarding was tied to personal performance, which was acting
against team centric thinking and the agile model in general [P3, P6, P49].
One case reported even on a practice of tracking failures down to individuals
[P40].

\subsubsection{Integrating requirements management and quality assurance}

Traditionally development is preceeded by requirements engineering and succeeded
by quality assurance (QA) activities. Agile methods suggest that essential parts
of these activities would be performed within the development team. However,
when developming large systems the effort required for managing requirements and
QA may be too heavy for individual teams to handle.

% Requirements engineering is difficult

In agile methods requirements engineering is usually closely tied to the
development team. However, large development projects will demand additional
high level requirements management. This is apparent in cases where complexity
of the product will require additional layers of product management [P45],
requirements are created by several stakehoder groups and development teams can
not be in contact with all of them [P39], or teams do not have access to
stakeholders due to a distributed seetting [P24]. Agile methods need to be
extended to include a process on how to get large requirements into team
backlogs [P31].

% As an integral part of the software development lifecycle, requirements
% management will be affected by the new agile process. In large organizations
% product and requirements management was initially done in a traditional way, but
% it was contrasted by the agile process that required iteratively supplying small
% requirements to development [P51].

Several cases highlighted that much learning was needed to master the new
process of creating user stories, both on product management and development
levels [P27, P30, P33, P37]. There were problems such as ambiguity in
requirements [P5, P11], and effort estimation for stories was hard to do well
[P27, P37].

Requirements were in some cases first defined on a high level in marketing
requirements documents (MRD) [P1, P31] or functional specifications [P33].
These high level requirements needed to be elaborated to be used by agile
development teams. The requirements breakdown in it self was found to be
difficult [P1, P33], and also when to do the breakdown and to which detail level
[P17].
Product managers and business analysts were struggling in creating high level
requirements [P33, P37], and teams were struggling in breaking them down to an
estimable size [P33]. Smits and Rilliet [P48] describe how high level product
management delivered requirements in too large chunks resulting in development
teams spending large amounts of time on defining suitaby sized stories.
% ?
In one case the teams were overwhelmed by reprioritization and requirements
bartering when stakeholders learned that requirements could be rearranged
flexibly [P31].

% Long term planning

High level requirements work and estimation was problematic as there was a gap
between short and long term planning [P9, P10]. Typical agile backlogs would
give only short term visibility, so organization specific practices for long
term planning had to be created [P26, P31]. Several cases described that care
had to be taken to avoid long term planning becoming a prescriptive practice. In
order to preserve the agility of development a schedule driven approach was
avoided [P10], sensitivity was used when considering setting milestones [P13],
and striving for more accurate estimates was not let to become an excuse for
gathering requirements up front [P38].

% Problems with enterprise or external QA

In an enterprise setting quality assurance tasks will require effort beyond the
general agile convention of doing testing within a crossfunctional team.
Similarly as with requirements engineering, the agile model may need to be
extended to accomodate additional testing activities [P31]. Organizations may
have existing separate QA teams that must be coordinated to work with
development teams [P17, P30, P48]. Integrating enterprise scale QA will pose
some additional challenges to agile methods.

Full quality assurance of a system requires special testing such as performance,
load, and memory testing, but agile methods lacks focus on them [P17]. These
testing activities can not be done within boundaries of user stories, and
require more resources than teams can spare [P31]. When testing tasks overlap
team boundaries it is sensible having a separate teams for the tasks, but
coordination practice between specialized QA teams and development teams needs
to be defined [P28, P48].

Some cases reported on problems with lack of automated testing. For a large
system the lack of automated tests caused excess testing work and late discovery
of defects [P10, P11, P17].

Still one challenge in QA was the relationship to requirements engineering. The
QA functions were struggling if there was problems such as ambiguity or
ineffective breakdown of large requirements [P48]. In a waterfall model an
extended time was reserved for QA at the end of projects, which allowed to
retroactively resolve ambiguities in requirements [P5]. However, ambiguity
became an impediment to QA when development worked in short increments and
there was no extended period for stabilization [P5].

\subsubsection{Other challenges in transformation}

There were still some other challenges in the transformation that do not fit in
the categories presented above. Two particular problems are worth mentioning.
Firstly, there were certain factors that pulled people back to the old ways of
working, and secondly, excessive enthusiasm in agile had negative effects.

% Falling back to old model

In several cases challenges in the transformation resulted in people falling
back to the old model of working. In some cases it was only a temporary struggle
while learning agile practices, but in other cases the old model displaced
agile.
Development work has to go on during the transformation but there will be new
things to learn for the team. Stress caused by the combination of schedule
pressure and much change at once will pull people back to the old way of working
[P11, P18, P38].
Schatz and Abdelshafi tell how even subtle trouble may risk the transformation,
as people will always look for reasons to revert to familiar behavior [P44].
Teams without adequate training were struggling with applying agile practices
correctly, and the challenge the new practices posed made people to abandon them
and return to the ways they know [P6, P49].

In some cases agile was seen as an overhead, and was therefore abandoned. For
instance, as new practices were being introduced there was a decrease
performance, and when teams realized that the benefits are not immediate they
started reverting to the old model [P21].
Evans [P12] describes that new senior appointments changed the management to be
less favorable towards agile, and a waterfall development model started to
return.

% Agile zealots

A phenomena that troubled some organizations was overenthusiasm towards agile
methods. In several cases there was mentionings about change leaders or teams
becoming agile zealots. For instance, Evans [12] describes how some members of
the agile community were becoming too evangelic, which made people take sides
for or against agile development. In another case team members started out with
an overly eager attitude, but when agile did not deliver improvements the
interest faded and started to go back to the old process [P20]. Attempting to
implement agile too strictly may cause conflicts, and especially in a large
organization the change leaders need to maintain a collaborative attitude
towards various groups in the surrounding organization [P6]. Introducing agile
methods does not guarantee success, and therefore they should not be followed
blindly [P4].

% When a release schedule with fixed deadlines that business required was combined
% to agile managers started to behave authoritatively and the self organization of
% agile started to break down [P31]. % (Skeptics used this as a sign of that agile can not work)

% This??
% % People do not speak their minds
% In one case a passive-aggressive organization culture was described as a
% challenge. Ranganath describes how people did not speak their minds in meetings,
% thus creating a false sense of agreement when talking about how the organization
% should develop. Because of this resistance and concerns were left growing, and
% acted against achieving a successful transformation. To to avoid these issues
% change leaders worked with team members on an individual level to win confidence 
% and make their voices heard [P40].

\clearpage

\subsection{Success factors in transformation}

Some success factors ...

%%%%%%%%%%%%%%%%%%%%%%%%%%%

% ---------------------------------------------------------------------

% ---------------------------------------------------------------------

% ---------------------------------------------------------------------
\section{Discussion}
\label{sec:discussion}

* Some success factors and challenges relate

* Success factors and challenges may be complements, and may be presented as
  complements

* What is special in large scale?

* What is similar in organizations of all scales?

* Transformation and scaling up are connected concepts. Although this research
  focuses on transformation process only, it can not be completely separated
  from discussion on scaling up, or on large scale agile implementation in
  general.

* Transformation of an organization to use agile can be seen as transforming
  also the agile method to suit the organization.
  --> A large scale agile model may require tailoring and trial over time

* It can be argued that any development effort with hundreds of teams will be
  difficult to coordinate

* Even though most of the primary studies are experience reports and subject to
  bias, this study can draw valuable conclusions. --> If a matter is being
  talked about it must be of some significance.

* Have flexibility for different teams in the agile model (agile found
  hard [P41]) vs. conformity to a single model.

* The factors that are listed are the ones to discuss -- there may be other
  factors affecting transformation, but this study provides evidence that these
  are the most important ones. We included only challenges that are discussed in
  more than studies, which proves that some filtering has been done.

* System development life cycle ?? -- There are inevitably other areas than
  development of increments, even in an agile process

% ---------------------------------------------------------------------
\section{Limitations and future research}
\label{sec:conclusion}

* Compare to change leadership research, and organizational research. Which
  challenges are typical to any organization, which are specific to software
  development organizations.

* Compare how recommendations in large scale application models of agile relate
  to the challenges and success factors reported in the transformaiton proecss.

% ---------------------------------------------------------------------
% -------------- BIBLIOGRAPHY -----------------------------------------
% ---------------------------------------------------------------------

\bibliographystyle{elsarticle-harv}
%\bibliographystylegen{plain}

\bibliography{sources_background,sources_primary,sources_excluded}

\end{document}